---
title: Speed up and parallelize
name: speed-parallel
description: Threads, grain size, and method choices for big data.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
params:
  family: green
  base_size: 13
  content_width: 72
vignette: |
  %\VignetteIndexEntry{Speed up and parallelize} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
css: albers.css
resource_files:
- albers.css
- albers.js
includes:
  in_header: |-
    <script src="albers.js"></script>
    <script>document.addEventListener('DOMContentLoaded',()=>document.body.classList.add('palette-red'));</script>

---

```{r setup, include=FALSE}
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.align = "center", fig.retina = 2,
  out.width = "100%", fig.width = 7, fig.asp = 0.618,
  message = FALSE, warning = FALSE
)
set.seed(123); options(pillar.sigfig = 7, width = 80)
suppressWarnings(suppressPackageStartupMessages({
  library(neurocluster)
  library(neuroim2)
}))
if (requireNamespace("albersdown", quietly = TRUE) && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::theme_set(albersdown::theme_albers(params$family, base_size = params$base_size))
}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```

```{r albers_inject, echo=FALSE, results='asis'}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```

# Threads {#threads}
```r
RcppParallel::setThreadOptions(numThreads = 2)
```

# Method choices {#method-choices}
- Large N: `slice_msf` (slice-wise MSF) or `flash3d` (hash/DCT); both avoid global iteration over all voxels.
- Preserve K: `cluster4d_slic()` with `preserve_k = TRUE`.
- Iterative refinement: `supervoxels()` performs multiple reassignment iterations; uses RcppParallel for updates.

# Grain size {#grain-size}
- For `supervoxels`, a practical start is `grain_size = max(100, nvox / (threads * 10))`. Smaller values improve balancing at some overhead.

# Memory tips {#memory-tips}
- `snic()` is single-pass with low memory overhead.
- Reduce time points or compress features if memory-bound.

# Toy runtime illustration (small N) {#fig-runtime}
Illustrative timing on a small, deterministic dataset (3-band synthetic). Absolute times are not representative of large datasets but show relative ordering on the same volume.

```{r fig-runtime, fig.cap='Illustrative runtimes on a 3-band synthetic (small N).', fig.alt='Bar plot of elapsed seconds for several methods on tiny data.'}
toy <- make_block_synthetic(dims = c(14, 14, 2), ntime = 40, noise = 0.1, seed = 321)

run_and_time <- function(tag, fn) {
  t <- system.time(fn())
  unname(t["elapsed"])
}

elapsed <- c(
  snic = run_and_time("snic", function() snic(toy$vec, toy$mask, K = 3, compactness = 3, max_iter = 3)),
  slice_msf = run_and_time("slice_msf", function() slice_msf(toy$vec, toy$mask,
                                                            target_k_global = 3,
                                                            r = 8, min_size = 6,
                                                            compactness = 3,
                                                            num_runs = 1,
                                                            stitch_z = FALSE)),
  supervoxels = run_and_time("supervoxels", function() supervoxels(toy$vec, toy$mask,
                                                                  K = 3,
                                                                  alpha = 0.6,
                                                                  connectivity = 6,
                                                                  iterations = 5)),
  slic = run_and_time("slic", function() cluster4d(toy$vec, toy$mask,
                                                  n_clusters = 3,
                                                  method = "slic",
                                                  spatial_weight = 0.5,
                                                  connectivity = 26,
                                                  max_iterations = 5,
                                                  preserve_k = TRUE))
)

barplot(elapsed, ylab = "seconds", main = "Toy runtimes (small N)")
```
includes:
  in_header: |-
    <link rel="stylesheet" href="albers.css" />
    <script src="albers.js"></script>
