---
title: Legacy APIs and backward compatibility
name: legacy-apis
description: Map old interfaces to the unified cluster4d entry point.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
params:
  family: ochre
  base_size: 13
  content_width: 72
vignette: |
  %\VignetteIndexEntry{Legacy APIs and backward compatibility} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
css: albers.css
resource_files:
- albers.css
- albers.js
includes:
  in_header: |-
    <script src="albers.js"></script>
    <script>document.addEventListener('DOMContentLoaded',()=>document.body.classList.add('palette-red'));</script>

---

# Map old to new {#map}
```{r setup, include=FALSE}
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.align = "center", fig.retina = 2,
  out.width = "100%", fig.width = 7, fig.asp = 0.618,
  message = FALSE, warning = FALSE
)
set.seed(123); options(pillar.sigfig = 7, width = 80)
if (requireNamespace("albersdown", quietly = TRUE) && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::theme_set(albersdown::theme_albers(params$family, base_size = params$base_size))
}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```

```{r albers_inject, echo=FALSE, results='asis'}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```
- `supervoxels(vec, mask, K, alpha, ...)` → `cluster4d(vec, mask, n_clusters=K, method="supervoxels", spatial_weight=1-alpha, ...)`
- `snic(vec, mask, K, compactness, ...)` → `cluster4d(..., method="snic", spatial_weight=compactness/10, ...)`
- `slic4d_supervoxels(...)` → `cluster4d(..., method="slic", ...)`
- `slice_msf(...)` → `cluster4d(..., method="slice_msf", ...)`
- `supervoxels_flash3d(...)` → `cluster4d(..., method="flash3d", ...)`

# Examples {#examples}
```r
# Old
# res <- supervoxels(vec, mask, K = 100, alpha = 0.3)

# New
# res <- cluster4d(vec, mask, n_clusters = 100, method = "supervoxels", spatial_weight = 0.7)
```

# Deprecations {#deprecations}
- Prefer `cluster4d()` for consistency and shared validation.
includes:
  in_header: |-
    <link rel="stylesheet" href="albers.css" />
    <script src="albers.js"></script>
