---
title: Method deep dives
name: method-deep-dives
description: Supervoxels, SNIC, SLIC, slice-MSF, and FLASH-3D—ideas and tradeoffs.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
params:
  family: teal
  base_size: 13
  content_width: 72
vignette: |
  %\VignetteIndexEntry{Method deep dives} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
css: albers.css
resource_files:
- albers.css
- albers.js
includes:
  in_header: |-
    <script src="albers.js"></script>
    <script>document.addEventListener('DOMContentLoaded',()=>document.body.classList.add('palette-red'));</script>

---

# Supervoxels {#supervoxels}
```{r setup, include=FALSE}
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.align = "center", fig.retina = 2,
  out.width = "100%", fig.width = 7, fig.asp = 0.618,
  message = FALSE, warning = FALSE
)
set.seed(123); options(pillar.sigfig = 7, width = 80)
if (requireNamespace("albersdown", quietly = TRUE) && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::theme_set(albersdown::theme_albers(params$family, base_size = params$base_size))
}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```

```{r albers_inject, echo=FALSE, results='asis'}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```
- Iterative, heat-kernel similarities; parallelized updates available.

# SNIC {#snic}
- Non-iterative; sequential priority queue; connected components by design.

# SLIC {#slic}
- Iterative with local windows; preserves K; good speed/quality balance.

# Slice-MSF {#slice-msf}
- Slice-wise minimum spanning forest; supports consensus and stitching.

# FLASH-3D {#flash3d}
- DCT-based hashing/compression; fast approximation; temporal smoothing via `lambda_t`.
includes:
  in_header: |-
    <link rel="stylesheet" href="albers.css" />
    <script src="albers.js"></script>
