---
title: API overview
name: api-overview
description: Group exported functions by task and link to details.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
params:
  family: red
  base_size: 13
  content_width: 72
vignette: |
  %\VignetteIndexEntry{API overview} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
css: albers.css
resource_files:
- albers.css
- albers.js
includes:
  in_header: |-
    <script src="albers.js"></script>
    <script>document.addEventListener('DOMContentLoaded',()=>document.body.classList.add('palette-red'));</script>

---

```{r setup, include=FALSE}
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.align = "center", fig.retina = 2,
  out.width = "100%", fig.width = 7, fig.asp = 0.618,
  message = FALSE, warning = FALSE
)
set.seed(123); options(pillar.sigfig = 7, width = 80)
if (requireNamespace("albersdown", quietly = TRUE) && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::theme_set(albersdown::theme_albers(params$family, base_size = params$base_size))
  css_file <- system.file("pkgdown/assets/albers.css", package = "albersdown")
  if (file.exists(css_file)) {
    cat(sprintf("<style>\n%s\n</style>", paste(readLines(css_file, warn = FALSE), collapse = "\n")))
  }
}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```

```{r albers_inject, echo=FALSE, results='asis'}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```

# Interfaces {#interfaces}
- Unified entry: `cluster4d()`
- Wrappers: `cluster4d_supervoxels`, `cluster4d_snic`, `cluster4d_slic`, `cluster4d_slice_msf`, `cluster4d_flash3d`

# Algorithms {#algorithms}
- Legacy cores: `supervoxels`, `snic`, `slic4d_supervoxels`, `slice_msf`, `supervoxels_flash3d`

# Utilities {#utilities}
- Validation & comparison: `validate_cluster4d`, `compare_cluster4d`
- Helpers: `compute_centroids`, `knn_shrink`, `spatial_gradient`, `tesselate`

# Variants {#variants}
- Surfaces: `supervoxel_cluster_surface`
- Time: `supervoxel_cluster_time`
includes:
  in_header: |-
    <link rel="stylesheet" href="albers.css" />
    <script src="albers.js"></script>
