---
title: Objects and results
name: objects-results
description: cluster4d_result structure, S3 methods, and neuroim2 objects.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
params:
  family: lapis
  base_size: 13
  content_width: 72
vignette: |
  %\VignetteIndexEntry{Objects and results} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
css: albers.css
includes:
  in_header: |-
    <script src="albers.js"></script>
    <script>document.addEventListener('DOMContentLoaded',()=>document.body.classList.add('palette-red'));</script>
resource_files:
- albers.css
- albers.js

---

# cluster4d_result {#cluster4d-result}
```{r setup, include=FALSE}
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.align = "center", fig.retina = 2,
  out.width = "100%", fig.width = 7, fig.asp = 0.618,
  message = FALSE, warning = FALSE
)
set.seed(123); options(pillar.sigfig = 7, width = 80)
if (requireNamespace("albersdown", quietly = TRUE) && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::theme_set(albersdown::theme_albers(params$family, base_size = params$base_size))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```

```{r albers_inject, echo=FALSE, results='asis'}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```
Fields: `clusvol`, `cluster`, `centers`, `coord_centers`, `n_clusters`, `method`, `parameters`, `metadata`.

# Print/plot/summary {#plot-print-summary}
- `print(x)`: method, counts, cluster size stats
- `summary(x)`: parameters, cluster size distribution, ranges
- `plot(x)`: axial/sagittal/coronal slices

# neuroim2 objects {#neuroim2-objects}
- `NeuroVol`, `NeuroVec`, `ClusteredNeuroVol`, helpers (`series`, `index_to_coord`, `spacing`)
