---
title: "Choose the right neurocluster interface"
name: api-overview
description: Start with cluster4d, then move to validation, comparison, or direct method APIs.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
    css: albers.css
    includes:
      in_header: albers-header.html
params:
  family: red
  base_size: 13
  content_width: 72
  preset: homage
resource_files:
- albers.css
- albers.js
- albers-header.html
- fonts
vignette: |
  %\VignetteIndexEntry{Choose the right neurocluster interface}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE,
  fig.align = "center", fig.retina = 2, out.width = "100%"
)
if (requireNamespace("albersdown", quietly = TRUE) &&
    requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::theme_set(albersdown::theme_albers(
    family = params$family, preset = params$preset
  ))
}
library(neurocluster)
```

```{r albers-classes, echo=FALSE, results='asis'}
cat(sprintf(
  paste0('<script>document.addEventListener("DOMContentLoaded",function(){',
         'document.body.classList.add("palette-%s","preset-%s");',
         '});</script><style>:root{--content:%sch}</style>'),
  params$family, params$preset, params$content_width
))
```

`neurocluster` has one main workflow and several specialist surfaces. If you
are starting an analysis, use `cluster4d()`: it gives every method the same
input checks and the same result contract. Move to a direct method function
only when you need controls that the unified interface intentionally does not
standardize.

## What is the shortest successful path?

| Task | Start here | What comes back |
|---|---|---|
| Fit one spatial clustering | `cluster4d()` | A validated-shape `cluster4d_result` |
| Check a fitted result against its inputs | `validate_cluster4d()` | Errors, warnings, and a compact summary |
| Compare compatible fits | `compare_cluster4d()` | Explicit size, dispersion, coherence, or agreement estimands |
| Choose a defensible setting | `suggest_cluster4d_params()` | A starting point to validate on your own data |
| Create a reproducible example | `generate_synthetic_volume()` | Data, mask, and known latent labels |

The complete first-use journey is in [Getting started](getting-started.html).
Parameter effects and validation are covered by
[Choose parameters](choose-parameters.html) and
[Validate and compare clusterings](validate-compare.html).

## Which methods share the unified interface?

```{r method-list}
methods <- eval(formals(cluster4d)$method)
paste(methods, collapse = ", ")
```

```{r method-list-check, include=FALSE}
stopifnot(
  length(methods) == 13L,
  identical(
    methods,
    c("supervoxels", "snic", "slic", "corr_slic", "brs_slic",
      "slice_msf", "flash3d", "g3s", "rena", "rena_plus", "mcl",
      "acsc", "commute")
  )
)
```

The list above is read from the installed function, so it cannot silently
drift from the current API. Method names select implementations; they are not
a performance ranking or endorsement. In particular, `slice_msf` remains
experimental, is excluded from automatic recommendations, and should only be
used for explicit evaluation under its [documented contract](slice-msf.html).
See [Compare clustering methods](compare-methods.html) for a scoped comparison
and [Method deep dives](method-deep-dives.html) for algorithmic distinctions.

## When should you use a direct function?

Typed method-specific wrappers expose native controls while retaining the
standard result contract. Examples include `cluster4d_snic()`,
`cluster4d_slic()`, `cluster4d_g3s()`, `cluster4d_rena()`,
`cluster4d_mcl()`, and `cluster4d_acsc()`.

```{r direct-export-check, include=FALSE}
typed_wrappers <- c(
  "cluster4d_supervoxels", "cluster4d_snic", "cluster4d_slic",
  "cluster4d_slice_msf", "cluster4d_flash3d", "cluster4d_g3s",
  "cluster4d_rena", "cluster4d_rena_plus", "cluster4d_mcl",
  "cluster4d_acsc", "cluster4d_commute"
)
stopifnot(all(typed_wrappers %in% getNamespaceExports("neurocluster")))
```

The current `cluster4d()` documentation classifies `supervoxels()`, `snic()`,
`slic4d_supervoxels()`, `slice_msf()`, and `supervoxels_flash3d()` as legacy,
deprecated entry points. They remain exported for compatibility, but new code
should use `cluster4d()` or a typed `cluster4d_*()` wrapper. Arguments and
defaults need not match across lifecycle generations. The mapping guide is
[Direct method APIs and cluster4d](legacy-apis.html).

## What should you inspect after fitting?

Use the result rather than reconstructing state from method internals:

```r
fit$labels          # labels in included-mask voxel order
fit$actual_k        # number of labels actually produced
fit$centers         # cluster means in the original feature space
fit$coord_centers   # cluster centroids in physical millimetres
fit$parameters      # recorded fitting parameters
fit$provenance      # geometry, mask, feature, and label provenance
```

[Read clustering results](objects-results.html) explains these fields and shows
the validation receipt that should accompany a saved or exported fit.
