---
title: Compare clustering methods
name: compare-methods
description: Compare valid partitions using explicit, shared estimands.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
    css: albers.css
    includes:
      in_header: albers-header.html
params:
  family: red
  base_size: 13
  content_width: 72
  preset: homage
resource_files:
- albers.css
- albers.js
- albers-header.html
- fonts
vignette: |
  %\VignetteIndexEntry{Compare clustering methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
if (requireNamespace("albersdown", quietly = TRUE)) albersdown::albers_register_fonts()
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.align = "center", fig.retina = 2,
  out.width = "100%", fig.width = 9, fig.asp = 0.56,
  message = FALSE, warning = FALSE
)
options(width = 90)
suppressPackageStartupMessages({
  library(neurocluster)
  library(neuroim2)
})
parcel_colors <- c("#0072B2", "#E69F00", "#009E73", "#CC79A7",
                   "#56B4E9", "#D55E00", "#F0E442", "#000000")
method_colors <- c(ReNA = "#0072B2", G3S = "#E69F00", SLIC = "#009E73")
align_labels <- function(labels, truth) {
  ids <- sort(unique(labels))
  truth_ids <- sort(unique(truth))
  stopifnot(length(ids) == length(truth_ids))
  overlap <- table(factor(labels, levels = ids),
                   factor(truth, levels = truth_ids))
  map <- truth_ids[as.integer(clue::solve_LSAP(overlap, maximum = TRUE))]
  aligned <- unname(map[match(labels, ids)])
  stopifnot(length(unique(aligned)) == length(ids))
  aligned
}
draw_slice <- function(labels, dims, z, title) {
  slice <- array(labels, dims)[, , z]
  image(t(slice[, ncol(slice):1]), col = parcel_colors, axes = FALSE,
        useRaster = TRUE, asp = 1, main = title)
  box(col = "#4A4A4A")
}
```

```{r albers-classes, echo=FALSE, results='asis'}
cat(sprintf(paste0(
  '<script>document.addEventListener("DOMContentLoaded",function(){',
  'document.body.classList.add("palette-%s","preset-%s");',
  '});</script><style>:root{--content:%sch}</style>'
), params$family, params$preset, params$content_width))
```

# What makes a comparison fair? {#fair-comparison}

Method comparison only makes sense when every result uses the same included
voxels and physical geometry. It also needs named estimands: parcel count and
size, spatial dispersion in millimetres, or temporal coherence from the original
voxel time series. `compare_cluster4d()` verifies the shared provenance before
computing those quantities.

We use one deterministic volume with known labels. This gives us an additional
simulation-only estimand—agreement with the known partition—without pretending
that synthetic recovery proves performance on real data.

```{r comparison-data}
syn <- generate_synthetic_volume(
  scenario = "gaussian_blobs", dims = c(16, 16, 8),
  n_clusters = 4, n_time = 40, noise_sd = 0.08, seed = 42
)
```

# How do you hold inputs constant? {#fit-methods}

Here ReNA, G3S, and SLIC all receive K = 4 and 26-neighbor connectivity. Their
other defaults remain method-specific, so this is a controlled workflow example,
not a tuned leaderboard. `preserve_k = TRUE` asks SLIC to repair its final
partition to the requested count.

```{r run-methods}
rena_fit <- cluster4d(
  syn$vec, syn$mask, 4, method = "rena",
  connectivity = 26, parallel = FALSE
)
g3s_fit <- cluster4d(
  syn$vec, syn$mask, 4, method = "g3s",
  connectivity = 26, parallel = FALSE
)
slic_fit <- cluster4d(
  syn$vec, syn$mask, 4, method = "slic",
  connectivity = 26, parallel = FALSE, preserve_k = TRUE
)
```

```{r check-methods, include=FALSE}
fits <- list(ReNA = rena_fit, G3S = g3s_fit, SLIC = slic_fit)
checks <- lapply(fits, validate_cluster4d, vec = syn$vec, mask = syn$mask)
truth_metrics <- lapply(fits, cluster_metrics, truth = syn$truth)
stopifnot(
  all(vapply(checks, `[[`, logical(1), "valid")),
  all(vapply(fits, `[[`, integer(1), "actual_k") == 4L),
  all(vapply(truth_metrics, function(x) is.finite(x$ari_truth), logical(1)))
)
feature_mat <- t(as.matrix(neuroim2::series(
  syn$vec, which(as.array(syn$mask) > 0)
)))
```

```{r method-slices, echo=FALSE, fig.cap='The known middle axial slice beside three fitted partitions. A one-to-one maximum-overlap assignment aligns the four arbitrary fitted IDs to the four known colors without merging predicted parcels; this display-only relabeling does not change any metric.', fig.alt='Four axial grid panels labeled known, ReNA, G3S, and SLIC. Every predicted parcel has a distinct color; ReNA matches the known regions while G3S and SLIC show different boundaries.'}
old_par <- par(mfrow = c(1, 4), mar = c(1, 1, 3, 1))
draw_slice(syn$truth, syn$dims, 4, "Known")
for (nm in names(fits)) {
  draw_slice(align_labels(fits[[nm]]$labels, syn$truth), syn$dims, 4, nm)
}
par(old_par)
```

# Which quantities are you comparing? {#estimands}

The call below evaluates all three methods on the same support. Lower spatial
RMS distance means voxels lie closer to their final parcel centroid. Temporal
pairwise correlation is the mean Pearson correlation over all unordered voxel
pairs within parcels; it is calculated from `feature_mat`, not from parcel
centers.

```{r compare-estimands}
comparison <- compare_cluster4d(
  fits,
  metrics = c("summary", "spatial_dispersion", "temporal_coherence"),
  feature_mat = feature_mat
)
comparison
```

```{r truth-estimands, echo=FALSE}
diagnostics <- data.frame(
  Method = names(fits),
  ARI_to_known = vapply(truth_metrics, `[[`, numeric(1), "ari_truth"),
  Spatial_RMS_mm = comparison$Spatial_RMS_Distance_mm,
  Temporal_correlation = comparison$Temporal_Pairwise_Correlation
)
knitr::kable(
  transform(diagnostics,
            ARI_to_known = round(ARI_to_known, 3),
            Spatial_RMS_mm = round(Spatial_RMS_mm, 2),
            Temporal_correlation = round(Temporal_correlation, 3)),
  caption = "Observed diagnostics on this synthetic fixture."
)
```

```{r check-comparison, include=FALSE}
stopifnot(
  identical(comparison$N_Clusters, rep(4L, 3)),
  all(is.finite(as.matrix(diagnostics[, -1]))),
  all(diagnostics$ARI_to_known >= -1 & diagnostics$ARI_to_known <= 1),
  all(diagnostics$Temporal_correlation >= -1 &
      diagnostics$Temporal_correlation <= 1)
)
```

```{r tradeoff-plot, echo=FALSE, fig.cap='Agreement with the known synthetic partition versus physical compactness. Up indicates greater agreement; left indicates smaller spatial RMS distance. No single direction optimizes both estimands here.', fig.alt='Scatter plot with spatial RMS distance on the horizontal axis and adjusted Rand index on the vertical axis. SLIC is leftmost and lowest, G3S is intermediate, and ReNA is upper right.'}
plot(
  diagnostics$Spatial_RMS_mm, diagnostics$ARI_to_known,
  xlab = "Spatial RMS distance (mm; lower is more compact)",
  ylab = "Adjusted Rand index to known partition (higher is closer)",
  pch = c(16, 17, 15), cex = 1.8,
  col = method_colors[diagnostics$Method], bty = "l",
  xlim = range(diagnostics$Spatial_RMS_mm) + c(-1.0, 1.0),
  ylim = c(0, 1.04)
)
text(diagnostics$Spatial_RMS_mm, diagnostics$ARI_to_known,
     labels = diagnostics$Method, pos = c(4, 2, 4), cex = 0.9)
abline(h = seq(0, 1, 0.25), col = "#D9D9D9", lty = 3)
```

On this fixture ReNA recovers the known partition, while SLIC yields the lowest
spatial dispersion. Those statements describe these evaluated outputs only.
Real method choice should be repeated across representative subjects, masks,
resolutions, and stability perturbations using estimands chosen before tuning.

# What should you compare next? {#next}

- Read [Choose parameters](choose-parameters.html) before tuning each method.
- Read [Validate and compare](validate-compare.html) for structural validation and pairwise partition agreement.
- Read [Benchmark methodology](benchmarks.html) for performance measurement rather than inferring speed here.
