---
title: "Direct method APIs and cluster4d"
name: legacy-apis
description: Understand what the unified interface maps and what remains method-specific.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
    css: albers.css
    includes:
      in_header: albers-header.html
params:
  family: red
  base_size: 13
  content_width: 72
  preset: homage
resource_files:
- albers.css
- albers.js
- albers-header.html
- fonts
vignette: |
  %\VignetteIndexEntry{Direct method APIs and cluster4d}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE,
  fig.align = "center", fig.retina = 2, out.width = "100%"
)
if (requireNamespace("albersdown", quietly = TRUE) &&
    requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::theme_set(albersdown::theme_albers(family = params$family, preset = params$preset))
}
library(neurocluster)
```

```{r albers-classes, echo=FALSE, results='asis'}
cat(sprintf(
  paste0('<script>document.addEventListener("DOMContentLoaded",function(){',
         'document.body.classList.add("palette-%s","preset-%s");',
         '});</script><style>:root{--content:%sch}</style>'),
  params$family, params$preset, params$content_width
))
```

The package keeps `cluster4d()`, typed `cluster4d_*()` wrappers, and older
method-native functions. The current `cluster4d()` documentation classifies
`supervoxels()`, `snic()`, `slic4d_supervoxels()`, `slice_msf()`, and
`supervoxels_flash3d()` as legacy/deprecated entry points. They remain exported
for compatibility, but new code should use the unified interface or its typed
wrappers. Equivalence should never be assumed unless mapped parameters are
identical.

`slice_msf()` and `cluster4d_slice_msf()` are additionally experimental. The
typed wrapper standardizes the result and requests exact K, but neither entry
point is currently recommended automatically or for production use. See the
[Slice-MSF evaluation contract](slice-msf.html) before evaluating either one.

## What does the unified interface map?

| Unified method | Legacy entry point | Preferred typed wrapper | Important mapping |
|---|---|---|---|
| `"supervoxels"` | `supervoxels()` | `cluster4d_supervoxels()` | `alpha = 1 - spatial_weight` |
| `"snic"` | `snic()` | `cluster4d_snic()` | `compactness = 10 * spatial_weight` |
| `"slic"` | `slic4d_supervoxels()` | `cluster4d_slic()` | `compactness = 20 * spatial_weight` |
| `"slice_msf"` (experimental) | `slice_msf()` | `cluster4d_slice_msf()` | natural versus exact K, connectivity, and FH scale are translated; positive `gamma` is rejected |
| `"flash3d"` | `supervoxels_flash3d()` | `cluster4d_flash3d()` | `lambda_s = spatial_weight`; temporal weight is complementary unless supplied |
| `"g3s"` | none (`g3s()` does not exist) | `cluster4d_g3s()` | `alpha = 1 - spatial_weight` |

Those statements describe the current wrappers. Defaults and method-specific
arguments can still differ, so record the actual `cluster4d()` or typed-wrapper
call—not an informal translation.

```{r lifecycle-check, include=FALSE}
exports <- getNamespaceExports("neurocluster")
legacy_exports <- c(
  "supervoxels", "snic", "slic4d_supervoxels", "slice_msf",
  "supervoxels_flash3d"
)
typed_exports <- c(
  "cluster4d_supervoxels", "cluster4d_snic", "cluster4d_slic",
  "cluster4d_slice_msf", "cluster4d_flash3d", "cluster4d_g3s"
)
stopifnot(
  all(legacy_exports %in% exports), all(typed_exports %in% exports),
  !"g3s" %in% exports
)
```

## Can a mapping be checked rather than trusted?

This small compatibility example checks the exact legacy SNIC mapping on one
generated volume. It explains an existing script; it is not a recommendation
to start new code with `snic()`.

```{r direct-data, echo=FALSE}
syn <- generate_synthetic_volume(
  scenario = "gaussian_blobs", dims = c(10, 10, 3),
  n_clusters = 4, n_time = 16, seed = 812
)
```

```{r direct-fit}
unified <- cluster4d(
  syn$vec, syn$mask, n_clusters = 4, method = "snic",
  spatial_weight = 0.4
)
direct <- snic(syn$vec, syn$mask, K = 4, compactness = 4)
```

```{r direct-check, include=FALSE}
vu <- validate_cluster4d(unified, syn$vec, syn$mask)
vd <- validate_cluster4d(direct, syn$vec, syn$mask)
stopifnot(vu$valid, vd$valid)
stopifnot(identical(unified$labels, direct$labels))
stopifnot(identical(unified$actual_k, direct$actual_k))
```

For this input and package version, the labels are identical. That is a check
of this mapping—not a blanket promise that every unified/direct pair is
bit-identical.

## Which interface should you publish?

Use `cluster4d()` when methods are compared under a shared contract. Use a
typed `cluster4d_*()` wrapper when the protocol depends on a native parameter.
Treat the legacy functions as compatibility surfaces. In every case, save the
result’s `parameters` and `provenance`, validate it against the original `vec`
and `mask`, and report the actual rather than merely requested cluster count.
See [Read clustering results](objects-results.html).
