---
title: "How the clustering methods differ"
name: method-deep-dives
description: Compare algorithmic work units and controls without turning them into universal rankings.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
    css: albers.css
    includes:
      in_header: albers-header.html
params:
  family: red
  base_size: 13
  content_width: 72
  preset: homage
resource_files:
- albers.css
- albers.js
- albers-header.html
- fonts
vignette: |
  %\VignetteIndexEntry{How the clustering methods differ}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE,
  fig.align = "center", fig.retina = 2, out.width = "100%",
  fig.width = 7, fig.asp = 0.58
)
if (requireNamespace("albersdown", quietly = TRUE) &&
    requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::theme_set(albersdown::theme_albers(family = params$family, preset = params$preset))
}
```

```{r albers-classes, echo=FALSE, results='asis'}
cat(sprintf(
  paste0('<script>document.addEventListener("DOMContentLoaded",function(){',
         'document.body.classList.add("palette-%s","preset-%s");',
         '});</script><style>:root{--content:%sch}</style>'),
  params$family, params$preset, params$content_width
))
```

All methods turn voxel-wise features and spatial support into parcels, but they
do not optimize the same objective. Choose a family because its assumptions
and controls match the analysis—not because an unscoped label calls it “fast”
or “high quality.”

## Where does each family spend its work?

```{r method-map-data, include=FALSE}
taxonomy <- data.frame(
  key = c("supervoxels", "snic", "slic", "corr_slic", "brs_slic",
          "slice_msf", "flash3d", "g3s", "rena", "rena_plus", "mcl",
          "acsc", "commute"),
  method = c("Supervoxels", "SNIC", "SLIC", "Corr-SLIC", "BRS-SLIC",
             "Slice-MSF", "FLASH-3D", "G3S", "ReNA", "ReNA+", "MCL",
             "ACSC", "Commute"),
  work_unit = c(
    "local heat kernels", "seeded local growth", "local assignment",
    "correlation embedding", "sketch plus exact boundary", "spatial graph",
    "compressed signatures", "gradient geodesic graph", "spatial graph",
    "spatial graph plus refinement", "weighted graph diffusion",
    "block community graph", "resistance embedding"
  ),
  update = c(
    "iterative", "single growth", "iterative", "iterative", "staged",
    "single construction", "iterative", "growth plus refinement",
    "agglomerative", "agglomerative", "iterative", "community plus refinement",
    "spectral then cluster"
  ),
  family = c("local", "local", "local", "embedded", "embedded", "graph",
             "embedded", "graph", "graph", "graph", "graph", "community",
             "spectral"),
  direct = c(
    "cluster4d_supervoxels", "cluster4d_snic", "cluster4d_slic", NA, NA,
    "cluster4d_slice_msf", "cluster4d_flash3d", "cluster4d_g3s",
    "cluster4d_rena", "cluster4d_rena_plus", "cluster4d_mcl",
    "cluster4d_acsc", "cluster4d_commute"
  ),
  stringsAsFactors = FALSE
)
method_choices <- eval(formals(neurocluster::cluster4d)$method)
package_exports <- getNamespaceExports("neurocluster")
stopifnot(
  identical(taxonomy$key, method_choices),
  nrow(taxonomy) == 13L,
  !anyDuplicated(taxonomy$key),
  all(stats::na.omit(taxonomy$direct) %in% package_exports),
  all(c("cluster4d", "validate_cluster4d", "compare_cluster4d") %in%
      package_exports)
)
family_order <- c("local", "embedded", "graph", "community", "spectral")
taxonomy$x <- match(taxonomy$family, family_order)
taxonomy$y <- ave(seq_len(nrow(taxonomy)), taxonomy$family, FUN = seq_along)
```

```{r taxonomy-table, echo=FALSE}
knitr::kable(
  taxonomy[c("method", "work_unit", "update")],
  col.names = c("Unified method", "Dominant work unit", "Update pattern"),
  caption = "All 13 choices accepted by the installed cluster4d() interface."
)
```

```{r method-map, echo=FALSE, fig.cap='All 13 unified methods grouped by their dominant representation. Vertical position only separates labels within a group; it is not a score.', fig.alt='Five grouped columns labeled local, embedded, graph, community, and spectral containing all thirteen method names.'}
palette_cb <- c(local = "#0072B2", embedded = "#D55E00", graph = "#009E73",
                community = "#CC79A7", spectral = "#E69F00")
plot(NA, xlim = c(0.5, 5.5), ylim = c(0.5, 5.2), axes = FALSE,
     xlab = "dominant representation", ylab = "")
axis(1, at = seq_along(family_order), labels = family_order)
abline(v = seq_along(family_order), col = "#E5E5E5", lty = 3)
points(taxonomy$x, taxonomy$y, pch = 21, cex = 2,
       bg = palette_cb[taxonomy$family], col = "white", lwd = 1.3)
text(taxonomy$x, taxonomy$y, labels = taxonomy$method,
     pos = 3, cex = 0.72)
box()
```

**What to notice.** The methods differ first in representation and update
strategy. The axes are categories, not claims about runtime, memory, or
accuracy.

## What is distinctive about each method?

- **SNIC** grows clusters from seeds with a priority queue in a non-iterative
  pass. Its native control is compactness.
- **SLIC** alternates local assignment and center updates. Its controls include
  compactness, iteration count, seeding, connectivity, and optional exact-K
  repair subject to mask topology.
- **Corr-SLIC** performs local SLIC updates in a compact correlation embedding;
  its approximation and exact-refinement controls must be reported.
- **BRS-SLIC** deliberately separates a coarse sketch stage from exact
  correlation refinement at parcel boundaries.
- **Supervoxels** iteratively combines feature and spatial heat-kernel
  similarities. It exposes bandwidths and convergence controls.
- **Slice-MSF** builds DCT sketches and segments a masked spatial graph; it can
  stitch across axial slices and combine seeded ensemble runs.
- **FLASH-3D** uses compact temporal signatures and iterative refinement. Hash
  length and DCT dimension are approximation controls, not generic quality
  scores.
- **G3S** uses functional gradients and geodesic growth followed by connected
  refinement.
- **ReNA** recursively agglomerates neighbouring regions on a spatial graph;
  **ReNA+** adds its own refinement path. Their weights are method-native rather
  than the unified spatial blend.
- **MCL** alternates expansion and inflation on a sparse weighted voxel graph;
  inflation controls granularity rather than a target-K objective.
- **ACSC** builds a block graph, applies community detection, and can refine
  boundaries using a chosen correlation definition.
- **Commute** embeds the voxel graph using effective-resistance geometry before
  clustering; its spectral representation has different scaling constraints
  from local assignment methods.

## What should drive the final choice?

Decide which properties are required: connected parcels, an exact target K,
cross-slice edges, a particular similarity definition, deterministic output,
or a memory ceiling. Then test candidate methods on the same mask and features
with defined estimands. [Compare clustering methods](compare-methods.html)
shows that workflow; [Benchmark with a reproducible receipt](benchmarks.html)
shows how to attach runtime evidence.
