---
title: "Reason about performance and memory"
name: performance-memory
description: Define what can be estimated before measuring method-specific runtime and memory.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
    css: albers.css
    includes:
      in_header: albers-header.html
params:
  family: red
  base_size: 13
  content_width: 72
  preset: homage
resource_files:
- albers.css
- albers.js
- albers-header.html
- fonts
vignette: |
  %\VignetteIndexEntry{Reason about performance and memory}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE,
  fig.align = "center", fig.retina = 2, out.width = "100%",
  fig.width = 7, fig.asp = 0.62
)
if (requireNamespace("albersdown", quietly = TRUE) &&
    requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::theme_set(albersdown::theme_albers(family = params$family, preset = params$preset))
}
```

```{r albers-classes, echo=FALSE, results='asis'}
cat(sprintf(
  paste0('<script>document.addEventListener("DOMContentLoaded",function(){',
         'document.body.classList.add("palette-%s","preset-%s");',
         '});</script><style>:root{--content:%sch}</style>'),
  params$family, params$preset, params$content_width
))
```

There is no universal fastest or smallest-memory method. Runtime depends on
mask topology, time points, requested K, parameter values, thread backend, and
hardware. Memory depends on implementation intermediates as well as the input.
This article gives one exact lower-level quantity you can estimate, then shows
how to produce evidence for your own workload.

## What can be estimated before a run?

If the masked feature matrix is materialized as dense R doubles with `N`
voxels and `T` features, its payload is exactly `8 * N * T` bytes. This is one
matrix, not peak process memory: coordinates, labels, graphs, sketches,
centroids, temporary copies, and allocator overhead are additional.

```{r memory-estimate}
dense_feature_mb <- function(n_voxels, n_features) {
  8 * n_voxels * n_features / 1e6
}
dense_feature_mb(100000, 200)
```

```{r memory-check, include=FALSE}
stopifnot(identical(dense_feature_mb(100000, 200), 160))
grid_n <- c(10000, 30000, 100000, 300000)
grid_t <- c(50, 100, 200, 400)
memory_grid <- outer(grid_n, grid_t, dense_feature_mb)
stopifnot(all(is.finite(memory_grid)), all(memory_grid > 0))
stopifnot(all(diff(memory_grid[, 3]) > 0), all(diff(memory_grid[3, ]) > 0))
```

```{r memory-figure, echo=FALSE, fig.cap='Payload size of one dense double feature matrix. This is not an estimate of peak process memory.', fig.alt='Heat map showing memory rising linearly with voxel count and number of features.'}
palette_cb <- colorRampPalette(c("#F0F0F0", "#56B4E9", "#0072B2"))(32)
image(
  x = seq_along(grid_n), y = seq_along(grid_t), z = log10(memory_grid),
  col = palette_cb, axes = FALSE, xlab = "masked voxels (N)",
  ylab = "features / time points (T)"
)
axis(1, at = seq_along(grid_n), labels = format(grid_n, big.mark = ","))
axis(2, at = seq_along(grid_t), labels = grid_t, las = 1)
text(
  rep(seq_along(grid_n), times = length(grid_t)),
  rep(seq_along(grid_t), each = length(grid_n)),
  labels = paste0(round(as.vector(memory_grid)), " MB")
)
box()
```

**What to notice.** Doubling either axis doubles this matrix’s payload. It does
not predict which algorithm will allocate copies or graph structures, so do
not turn the heat map into a method ranking.

## How should performance be measured?

Define the workload first: package version and commit, dimensions, included
voxels, features, K, method arguments, thread setting, machine, operating
system, and number of repetitions. Warm up compiled code, retain every repeat,
and report a distribution rather than the single best time. Check the fitted
result after each run so a fast failure or changed cluster count cannot win.

[Benchmark with a reproducible receipt](benchmarks.html) provides an executable
micro-benchmark and receipt template. [Parallel execution](speed-parallel.html)
explains thread controls without changing global thread state.
