---
title: Choose parameters for your data
name: choose-parameters
description: Use suggestions and tune K, spatial weighting, and connectivity.
output:
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 2
params:
  family: ochre
  base_size: 13
  content_width: 72
vignette: |
  %\VignetteIndexEntry{Choose parameters for your data} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
css: albers.css
resource_files:
- albers.css
- albers.js
includes:
  in_header: |-
    <script src="albers.js"></script>
    <script>document.addEventListener('DOMContentLoaded',()=>document.body.classList.add('palette-red'));</script>

---

```{r setup, include=FALSE}
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
if (requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(params$family))
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.align = "center", fig.retina = 2,
  out.width = "100%", fig.width = 7, fig.asp = 0.618,
  message = FALSE, warning = FALSE
)
set.seed(123); options(pillar.sigfig = 7, width = 80)
suppressWarnings(suppressPackageStartupMessages({
  library(neurocluster)
}))
if (requireNamespace("albersdown", quietly = TRUE) && requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::theme_set(albersdown::theme_albers(params$family, base_size = params$base_size))
}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```

```{r albers_inject, echo=FALSE, results='asis'}
if (file.exists("albers.css")) {
  cat(sprintf("<style>\n%s\n</style>", paste(readLines("albers.css", warn = FALSE), collapse = "\n")))
}
cat(sprintf('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("palette-%s");});</script>', params$family))
cat(sprintf('<style>:root{--content:%sch}</style>', params$content_width))
```

# Heuristics {#heuristics}
Rules of thumb: 1 cluster per ~100–500 voxels; start with `spatial_weight = 0.5`; use connectivity = 26 for smoother parcels.

# Suggestions {#suggestions}
```r
n_vox <- 50000
n_time <- 200
params <- suggest_cluster4d_params(n_vox, n_time, priority = "balanced")
params$recommended_method
params$n_clusters
```

# Tradeoffs {#tradeoffs}
- Quality: more iterations, supervoxels or SLIC, higher K
- Speed: slice_msf or flash3d, fewer iterations, lower K
- Memory: SNIC, slice_msf with fewer DCT components

# Recipes {#recipes}
- Whole-brain (2–3 mm): K=300–800, `spatial_weight=0.4–0.6`
- ROI: K=10–100, `spatial_weight=0.2–0.4`, connectivity=6
includes:
  in_header: |-
    <link rel="stylesheet" href="albers.css" />
    <script src="albers.js"></script>
